---
title: "music listening"
subtitle: before and after the pandemic
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
---

I've been working from home since mid-March 2020. This has afforded me the opportunity to listen to more of my collection of LPs and CDs than I might normally. But how much more?

I have been tracking my listening at home since November 2016, recording the artist, album, format, and date (along with a few other variables that don't figure into this analysis).

```{r setup}
# tidyverse
library(tidyverse)
library(lubridate)
library(readxl)
library(scales)

# utilities
library(janitor)
library(googlesheets4)
library(RcppRoll) 

# run `gs4_auth()` in the console to set google authorization
# - see https://googlesheets4.tidyverse.org/reference/gs4_auth.html for details


```

## read from google

https://www.r-bloggers.com/2021/12/how-to-connect-r-to-google-sheets-using-googlesheets4/

```{r google_read, message=FALSE}
# local copy
#df_playlist <- read_excel("music playlist.xlsx")


# read from google sheets

# test version
#music_playlist <- "https://docs.google.com/spreadsheets/d/1LJdGL1tASJYJ41Yvxku4k5_tYG5mV-8xcoH8e1-vWp8/edit?usp=sharing"

# full version
music_playlist <- "https://docs.google.com/spreadsheets/d/1kHBEMFEDwyszldNbNvD2A6Pt8KYoUACDZZ0usWvrruA/edit?usp=sharing"


#df_playlist <- read_sheet(music_playlist)
#df_playlist <- googlesheets4::read_sheet(music_playlist, 
#                          col_types = "cciiiiDccicccccccccccccccccccDdcclc"
#                          ) %>% 
#  rename("index" = tidyselect::contains("index"))
#
#tail(df_playlist)


# use the "range_speedread()" function to speed things up
# see https://googlesheets4.tidyverse.org/reference/range_speedread.html
df_playlist <- googlesheets4::range_speedread(music_playlist)
tail(df_playlist)

df_playlist <- googlesheets4::range_speedread(music_playlist,
                                              col_types = cols(
                                                disc_num = col_integer()
                                              ))

df_playlist <- df_playlist %>% 
  rename("index" = starts_with("index"))

tail(df_playlist)


```



```{r}

df_playlist_trim <- df_playlist %>% 
  mutate(disc_num = case_when(
    is.na(disc_num) ~ 1,
    TRUE ~ as.numeric(disc_num)
  )) %>% 
  select(index, 
         year, month, day,
         date_of_acquisition,
         format, disc_num,
         genre) %>% 
#  mutate(play_date = make_date(year, month, day)) %>% 
  mutate(play_date = as_date(glue::glue("{year}-{month}-{day}"))) %>% 
  filter(play_date > lubridate::as_date("2017-03-31"))

max(df_playlist_trim$play_date)
min(df_playlist_trim$play_date)

df_playlist_trim

df_playlist_trim %>% 
  group_by(date_of_acquisition) %>% 
  tally

df_playlist_trim %>% 
  group_by(play_date) %>% 
  tally()

df_play_summary <- df_playlist_trim %>% 
  group_by(play_date) %>%
  summarise(n = sum(disc_num))


df_play_summary <- df_play_summary %>%
  mutate(roll_avg = roll_mean(x = n, n = 14, fill = NA))

df_play_summary

ls.str(df_play_summary)

```

## plot trend

```{r plot_trend}


ggplot(df_play_summary, aes(x = play_date, y = roll_avg)) +
  geom_line() +
#  geom_smooth(span = 0.25)
## note: now there are >1000 observations, the function switches to
##  `method = "gam"` and `span = ` does not apply
  geom_smooth() 



```


```{r plot_trend_2023}

df_play_summary %>%
  filter(
    play_date > as_date("2023-01-01") 
#    &
#    play_date < as_date("2023-02-28")
    ) %>% 
  # plot
      ggplot(aes(x = play_date, y = roll_avg)) +
      geom_line() +
      geom_smooth(span = 0.75)


```

## ECM

```{r}

ecm <- df_playlist %>% 
  filter(str_detect(label, "ECM")) %>% 
  drop_na(year) %>% 
  mutate(play_date = make_date(year, month, day)) %>% 
  select(artist, 
         album, 
         format,
         play_date, 
         index)

#%>% 
#  filter(play_date > as_date("2017-03-31"))

ecm
```

### most played

#### artist

```{r}

ecm %>% 
  group_by(artist) %>% 
  tally() %>% 
  arrange(desc(n))

```

#### album

```{r}

ecm %>% 
  group_by(artist, album) %>% 
  tally() %>% 
  arrange(desc(n))

```


### only played once

```{r}

ecm_min <- ecm %>% 
#  filter(format == LP) %>% 
  group_by(artist, album) %>% 
  tally() %>% 
  filter(n == 1)
  
ecm_min

```


### most/least recently played


most recent

```{r}

ecm %>% 
  select(artist, 
         album, 
         format, 
         play_date, 
         index) %>% 
  arrange(desc(play_date), desc(index))

```

least recent (longest ago)

```{r}

ecm %>% 
  semi_join(ecm_min, by = c("artist", "album")) %>% 
  select(album, artist, play_date, index)

```



### ECM lps

#### longest-ago played LPs

These are the longest-ago played LPs, but may have been played more recently in another format

```{r ecm_longest_ago_lp}

ecm %>% 
  filter(format == "LP") %>% 
  group_by(artist, album) %>% 
  summarise(last_played = max(play_date)) %>% 
  ungroup() %>% 
  slice_min(order_by = last_played, n = 10)
  

```

#### longest-ago played LPs, full stop

Longest ago played LP, where it is only the LP that has been played

```{r}
ecm_last_played <- ecm %>% 
  group_by(artist, album) %>% 
  summarise(last_played = max(play_date)) %>% 
  ungroup()

# check unique
ecm_last_played %>% 
  group_by(artist, album) %>% 
  tally() %>% 
  filter(n > 1)

# create list of unique ECM LPs
ecm_lp <- ecm %>%
  filter(format == "LP") %>% 
  distinct(artist, album)

# join, slice with 10 least recently played
ecm_lp %>% 
  left_join(ecm_last_played, by = c("artist", "album")) %>% 
  slice_min(order_by = last_played, n = 20)
  

```



#### most recently played

```{r ecm_most_recent}
# option 1 - the entire list
ecm %>% 
  filter(format == "LP") %>% 
  arrange(desc(index))

# option 2: top 10
ecm %>% 
  filter(format == "LP") %>% 
  arrange(desc(index)) %>% 
  slice_head(n = 10)


```


## Pat Metheny

```{r pat_metheny}

metheny <- df_playlist %>% 
  filter(str_detect(artist, "Metheny")) %>% 
  drop_na(year) %>% 
  mutate(play_date = make_date(year, month, day)) %>% 
  select(artist, 
         album, 
         format,
         play_date)

metheny

```

### most played



```{r pat_metheny_most}

metheny %>% 
  group_by(album) %>% 
  tally() %>% 
  arrange(desc(n))

```


```{r}

metheny %>% 
  group_by(album) %>% 
  mutate(last_played = max(play_date)) %>% 
  ungroup() %>% 
  group_by(album, last_played) %>% 
  summarise(plays = n()) %>% 
  arrange(desc(plays), last_played)

```


## is this a replay of the same item?

This might be a challenge due to inconsistencies in data collection

```{r}
x <- tibble(
  value = c(1, 1, 2, 3, 4, 4),
  name = c("a", "b", "c", "d", "e", "e"))

x %>% 
  group_by(value, name) %>% 
  count()

```

LP and CD only

```{r}

df_playlist %>% 
  filter(format != "iTunes") %>%
  filter(format != "iMusic") %>% 
  filter(format != "iPhone") %>% 
  filter(format != "foobar2000") %>% 
  #
  group_by(artist, album, format) %>% 
  count() %>% 
  filter(n > 1) %>% 
  arrange(desc(n))
  


```

Everything


```{r}

df_playlist %>% 
  group_by(artist, album) %>% 
  count() %>% 
  filter(
    n > 1 &
    !is.na(artist)
    ) %>% 
  arrange(desc(n))
  


```


-30-